---
editor_options:
  chunk_output_type: console
---

# NOTES

-   This file is *not* included in the [source package](https://r-pkgs.org/structure.html#sec-source-package) because of the [`.nopurl` suffix in its
    filename](https://pkgpurl.rpkg.dev/reference/purl_rmd.html#-rmd-files-excluded-from-purling).

-   The chunks below have to be manually executed in order to regenerate the package data.

-   Although the datasets below are saved as ["internal data" in `R/sysdata.rda`](https://r-pkgs.org/data.html#sec-data-sysdata), they can still be exported and
    documented (by documenting the dataset's quoted name in the main `.Rmd` source file -- which only works when the dataset is also `@export`ed), something
    [not explicitly mentioned](https://coolbutuseless.github.io/2018/12/10/r-packages-internal-and-external-data/) in the book [R
    Packages](https://r-pkgs.org/data.html#sec-data-data). To do so, you first need to manually add the `export()` directive in the NAMESPACE file since
    roxygen2 won't add it automatically.

# Define data

## `data_abbrs`

```{r}
data_abbrs <-
  readr::read_csv(file = "data-raw/abbrs.csv",
                  col_types = "cc",
                  na = "") |>
  tibble::rowid_to_column() |>
  dplyr::group_by(abbreviation) |>
  dplyr::summarise(full_expressions = list(full_expression),
                   rowids = list(rowid),
                   .groups = "drop") |>
  dplyr::arrange(purrr::map_int(rowids,
                                dplyr::first)) |>
  dplyr::select(-rowids) |>
  dplyr::relocate(full_expressions)
```

## `data_md_snips`

```{r}
data_md_snips <-
  pal::toml_validate(input = "data-raw/md_snips.toml") |>
  pal::toml_read() |>
  purrr::imap(\(x, i) tibble::tibble(id = i,
                                     value = x$value)) |>
  purrr::list_rbind()
```

## `data_msgs`

```{r}
data_msgs <-
  pal::toml_validate(input = "data-raw/msgs.toml") |>
  pal::toml_read() |>
  purrr::imap(\(x, i) tibble::tibble(id = i,
                                     value = x$value)) |>
  purrr::list_rbind()
```

## `data_roxy_lbls`

```{r}
data_roxy_lbls <- tibble::tribble(
  ~type, ~name, ~label,
  NA, "data", "A [tibble][tibble::tbl_df].",
  NA, "data_cols", "A [tibble][tibble::tbl_df] with the columns {pal::prose_ls(cols, wrap = '`')}.",
  NA, "date", "A [date][base::Date].",
  NA, "dates", "A vector of [dates][base::Date].",
  NA, "datetime", "A [datetime][base::DateTimeClasses].",
  NA, "datetimes", "A vector of [datetimes][base::DateTimeClasses].",
  NA, "gt_obj", "A [`gt_tbl`][gt::gt] object.",
  NA, "path", "A [path][fs::fs_path].",
  NA, "paths", "A vector of [paths][fs::fs_path].",
  NA, "plotly_obj", "A [plotly object][plotly::plot_ly].",
  NA, "r_obj", "An \\R object.",
  NA, "response", "A [response object][httr::response].",
  NA, "strict_list", "A [strict list][xfun::strict_list()].",
  NA, "symbol", "A [symbol][base::name].",
  NA, "tabular_data", "A tabular data object like a [data frame][base::data.frame] or [tibble][tibble::tbl_df].",
  NA, "version_nr", "A [numeric version][numeric_version()].",
  NA, "opt_global_max_cache_age", paste0("Note that [on package load][base::ns-hooks], the cache will be cleared from entries exceeding a global maximum ",
                                         "timespan set by the \\R [option][base::options()] `{pkg}.global_max_cache_age` (defaults to ",
                                         "`{global_max_cache_age}` if unset)."),
  "description", "pkg_config", paste0("A [tibble][tibble::tbl_df] with metadata of all possible {pkg} package configuration options. See ",
                                      "[pal::pkg_config_val()] for more information."),
  "param", "cli_markup_support", "Supports cli's [inline markup][cli::inline-markup].",
  "param", "dyn_dots_support", "[Dynamic dots][rlang::dyn-dots] are supported.",
  "param", "tidy_select_support", "[Tidy selections][dplyr::dplyr_tidy_select] are supported.",
  "param", "eol", paste0("End of line (EOL) control character sequence. One of\n- `\"LF\"` for the line feed (LF) character (`\"\\n\"`). The standard on ",
                         "Unix and Unix-like systems (Linux, macOS, *BSD, etc.) and the **default**.\n- `\"CRLF\"` for the carriage return + line feed ",
                         "(CR+LF) character sequence (`\"\\r\\n\"`). The standard on Microsoft Windows, DOS and some other systems.\n- `\"CR\"` for the ",
                         "carriage return (CR) character (`\"\\r\"`). The standard on classic Mac OS and some other antiquated systems.\n- `\"LFCR\"` for ",
                         "the line feed + carriage return (LF+CR) character sequence (`\"\\n\\r\"`). The standard on RISC OS and some other exotic ",
                         "systems."),
  "param", "quiet", "Whether or not to suppress printing status output from internal processing.",
  "param", "start_date", "The begin of the period the data covers. A [date][base::Date] or a character scalar in the format `\"YYYY-MM-DD\"`.",
  "param", "end_date", "The end of the period the data covers. A [date][base::Date] or a character scalar in the format `\"YYYY-MM-DD\"`.",
  "param", "use_cache", "Return cached results if possible. If `FALSE`, results are always newly fetched regardless of `max_cache_age`.",
  "param", "max_cache_age", paste0("The duration after which cached results are refreshed (i.e. newly fetched). A valid ",
                                   "[lubridate duration][lubridate::as.duration]. Use `Inf` to disable cache expiry. Only relevant if `use_cache = TRUE`."),
  "title", "pkg_config", "{pkg} package configuration metadata"
)
```

# Write data

Save all the small data objects as a single internal file `R/sysdata.rda`. Note that when documenting them, they must be explicitly `@export`ed to be available
to package users.

```{r}
usethis::use_data(data_abbrs,
                  data_md_snips,
                  data_msgs,
                  data_roxy_lbls,
                  internal = TRUE,
                  overwrite = TRUE,
                  compress = "xz",
                  version = 3L)
```
