---
editor_options:
  chunk_output_type: console
---

# NOTES

-   This file is *not* included in the [source package](https://r-pkgs.org/structure.html#sec-source-package) because of the [`.nopurl` suffix in its
    filename](https://pkgpurl.rpkg.dev/reference/purl_rmd.html#-rmd-files-excluded-from-purling).

-   The chunks below have to be manually executed in order to regenerate the package data.

-   Although the datasets below are saved as ["internal data" in `R/sysdata.rda`](https://r-pkgs.org/data.html#sec-data-sysdata), they can still be exported and
    documented (by documenting the dataset's quoted name in the main `.Rmd` source file -- which only works when the dataset is also `@export`ed), something
    [not explicitly mentioned](https://coolbutuseless.github.io/2018/12/10/r-packages-internal-and-external-data/) in the book [R
    Packages](https://r-pkgs.org/data.html#sec-data-data). To do so, you first need to manually add the `export()` directive in the NAMESPACE file since
    roxygen2 won't add it automatically.

# Define data

## `roxy_tags_lbl`

```{r}
roxy_tags_lbl <- stringr::str_subset(string = roxygen2::tags_list(built_in = FALSE),
                                     pattern = pal::fuse_regex(paste0("^",
                                                                      c("aliases",
                                                                        "backref",
                                                                        "describeIn",
                                                                        "docType",
                                                                        "encoding",
                                                                        "family",
                                                                        "md",
                                                                        "method",
                                                                        "name",
                                                                        "noMd",
                                                                        "noRd",
                                                                        "note",
                                                                        "order",
                                                                        "rdname",
                                                                        "returns",
                                                                        "usage",
                                                                        "useDynLib"),
                                                                      "$"),
                                                               paste0("^",
                                                                      c("eval",
                                                                        "example",
                                                                        "export",
                                                                        "import",
                                                                        "include",
                                                                        "inherit",
                                                                        "raw",
                                                                        "template"))),
                                     negate = TRUE)
```

## `data_abbrs`

```{r}
data_abbrs <-
  readr::read_csv(file = "data-raw/abbrs.csv",
                  col_types = "cc",
                  na = "") |>
  tibble::rowid_to_column() |>
  dplyr::group_by(abbreviation) |>
  dplyr::summarise(full_expressions = list(full_expression),
                   rowids = list(rowid),
                   .groups = "drop") |>
  dplyr::arrange(purrr::map_int(rowids,
                                dplyr::first)) |>
  dplyr::select(-rowids) |>
  dplyr::relocate(full_expressions) |>
  assertr::assert(predicate = assertr::is_uniq,
                  "abbreviation")
```

## `data_md_snips`

```{r}
data_md_snips <-
  pal::toml_validate(input = "data-raw/md_snips.toml") |>
  pal::toml_read() |>
  purrr::imap(\(x, i) tibble::tibble(id = i,
                                     value = x$value)) |>
  purrr::list_rbind() |>
  assertr::assert(predicate = assertr::is_uniq,
                  "id")
```

## `data_roxy_lbls`

```{r}
data_roxy_lbls <- 
  pal::toml_validate(input = "data-raw/roxy_lbls.toml") |>
  pal::toml_read() |>
  purrr::imap(\(x, i) {
    
    checkmate::assert_subset(names(x$value),
                             choices = c("default",
                                         roxy_tags_lbl),
                             empty.ok = FALSE)
    
    if (!("default" %in% names(x$value))) {
      cli::cli_abort("No {.field default} value specified for roxy label with {.field id} {.val {i}} in {.file data-raw/roxy_lbls.toml}.")
    }
    
    purrr::map(names(x$value),
               \(key) tibble::tibble(id = i,
                                     type = key,
                                     value = x$value[[key]])) |>
      purrr::list_rbind()
  }) |>
  purrr::list_rbind() |>
  dplyr::arrange(id, type)

# assert id-type combos are unique
duplicated_data_roxy_lbls <-
  data_roxy_lbls |>
  dplyr::summarise(n = dplyr::n(),
                   .by = c("id", "type")) |>
  dplyr::filter(n > 1L)

if (nrow(duplicated_data_roxy_lbls) > 0L) {
  cli::cli_abort("{.var id}-{.var type} combos in {.file data-raw/roxy_lbls.toml} are not unique.")
}
```

# Write data

Save all the small data objects as a single internal file `R/sysdata.rda`. Note that when documenting them, they must be explicitly `@export`ed to be available
to package users.

```{r}
usethis::use_data(data_abbrs,
                  data_md_snips,
                  data_roxy_lbls,
                  roxy_tags_lbl,
                  internal = TRUE,
                  overwrite = TRUE,
                  compress = "xz",
                  version = 3L)
```
