---
output: pal::gitlab_document
---

# `r paste0(pal::desc_value(key = "Package"), ": ", pal::desc_value(key = "Title"))`

`r pal::desc_value(key = "Description")`

## R Markdown file snippets

The (R) Markdown file snippets can be used anywhere [(R) Markdown](https://rmarkdown.rstudio.com/) input is supported. For example, you can use them as input
to [knitr](https://yihui.org/knitr/)'s [`child` document option](https://yihui.org/knitr/options/#child-documents):

`r paste0("\x60\x60\x60\x60\n\x60\x60\x60", '{r, child = pkgsnip::rmd_snippet_path("installation-notice_dev-version_gitlab.Rmd")}\n', "\x60\x60\x60\n\x60\x60\x60\x60")`

When inserting snippets that require the name of the calling package like `"installation-notice_dev-version_gitlab.Rmd"`, you should either use
`pal::build_readme()` to build the `.Rmd` file or manually assign `pkg_metadata <- desc::desc_get(desc::desc_fields())` to the environment passed to
[`devtools::build_rmd()`](https://devtools.r-lib.org/reference/build_rmd.html) or its `...` argument respectively.

You can also use the snippets in [roxygen2](https://roxygen2.r-lib.org/)'s
[`@includeRmd` tag](https://roxygen2.r-lib.org/articles/rd.html#including-external--rmd-md-files):

```rd
#' @includeRmd `r '\x60r pkgsnip::rmd_snippet_path("rmd-package-notice.Rmd")\x60'`
```

For snippets that require the name of the calling package, just assign `pkg_metadata` (once) before inserting the snippet(s), e.g. like this:

```rd
#' `r "\x60r pkg_metadata <- desc::desc_get(desc::desc_fields())\x60"`
#'
#' @includeRmd `r '\x60r pkgsnip::rmd_snippet_path("installation-notice_dev-version.Rmd")\x60'`
```

### Included Snippets

Currently, the following R Markdown file snippets are included:

```{r, echo = FALSE, results = "asis"}
library(magrittr)

fs::path_package("snippets/",
                 package = pal::desc_value(key = "Package")) %>%
  fs::dir_ls(recurse = TRUE,
             type = "file") %>%
  fs::path_file() %>%
  paste0("- [`", ., "`](inst/snippets/", ., ")") %>%
  pal::cat_lines()
```

## roxygen2 tag labels

The [roxygen2](https://roxygen2.r-lib.org/) tag labels are specifically meant for [`@`-tag](https://roxygen2.r-lib.org/articles/rd.html)-based documentation
using [dynamic R code](https://roxygen2.r-lib.org/articles/rd-formatting.html#dynamic-r-code-1).

Insert them in your functions using inline R code, e.g. as follows

```r
`r paste0("#' @param use_cache \U0060r pkgsnip::param_label(\"use_cache\")\U0060")`
`r paste0("#' @param cache_lifespan \U0060r pkgsnip::param_label(\"cache_lifespan\")\U0060")`
`r paste0("#' @return \U0060r pkgsnip::return_label(\"data\")\U0060")`
```

which will result in

```r
#' @param use_cache `r pkgsnip::param_label("use_cache")`
#' @param cache_lifespan `r pkgsnip::param_label("cache_lifespan")`
#' @return `r pkgsnip::return_label("data")`
```

## R condition messages

These are meant to be used as message texts for base R's [`message()`](https://rdrr.io/r/base/message.html), [`warning()`](https://rdrr.io/r/base/warning.html)
and [`stop()`](https://rdrr.io/r/base/stop.html) functions or
[`rlang::inform()`, `rlang::warn()` and `rlang::abort()`](https://rlang.r-lib.org/reference/abort.html) from the tidyverse respectively.

A message can be retrieved using `pkgsnip::msg()`. For some of the messages, arguments need to be provided so they can be tailored to their specific
purpose. For example, the `pkg_required` message requires the `pkg` argument:

```r
pkgsnip::msg(name = "pkg_required",
             pkg = "some_pkg")
```

Which results in the following message text:

````{r, echo = FALSE, comment = ""}
pkgsnip::msg(name = "pkg_required",
             pkg = "some_pkg")
````

Currently, the following R condition messages are included:

```{r, echo = FALSE, results = "asis"}
pkgsnip::messages() %>%
  dplyr::mutate(dplyr::across(name,
                              ~ paste0("`", .x, "`"))) %>%
  pkgsnip:::add_args_col() %>%
  pal::pipe_table() %>%
  cat()
```

## Installation
```{r, install-snippet-dev, child = pkgsnip::rmd_snippet_path("installation-notice_dev-version_gitlab.Rmd")}
```

## Development

### R Markdown format
```{r, rmd-pkg-snippet, child = pkgsnip::rmd_snippet_path("rmd-package-notice.Rmd")}
```

### Coding style
```{r, coding-style-snippet, child = pkgsnip::rmd_snippet_path("coding-style-notice.Rmd")}
```

---
