---
output: pal::gitlab_document
---

# `r paste0(pal::desc_value(key = "Package"), ": ", pal::desc_value(key = "Title"))`

`r pal::desc_value(key = "Description")`

## (R) Markdown snippets

The (R) Markdown snippets included in this package can be used anywhere [(R) Markdown](https://rmarkdown.rstudio.com/) input is supported. For example, you can use them as input to [knitr](https://yihui.org/knitr/)'s [`child` document option](https://yihui.org/knitr/options/#child-documents):

```rmd
\```{r, child = pkgsnippets::snippet_path("installation-notice_dev-version_gitlab.Rmd")}
...
\```
```

Or you can use them in [roxygen2](https://roxygen2.r-lib.org/articles/rd.html#including-external--rmd-md-files)'s `@includeRmd` tag:

```rd
#' @includeRmd `pkgsnippets::snippet_path("literate-programming-notice.Rmd")`
```

### Included Snippets

Currently, the following snippets are included:

```{r, echo = FALSE, results = "asis"}
library(magrittr)

fs::path_package("snippets/",
                 package = pal::desc_value(key = "Package")) %>%
  fs::dir_ls(recurse = TRUE,
             type = "file") %>%
  fs::path_file() %>%
  paste0("- [`", ., "`](inst/snippets/", ., ")") %>%
  pal::cat_lines()
```

## roxygen2 tag labels

The [roxygen2](https://roxygen2.r-lib.org/) tag labels included in this package are specifically meant for [`@`-tag](https://roxygen2.r-lib.org/articles/rd.html)-based documentation using [dynamic R code](https://roxygen2.r-lib.org/articles/rd-formatting.html#dynamic-r-code-1).

Insert them in your functions using inline R code, e.g. as follows

```r
`r paste0("#' @param use_cache \U0060r pkgsnippets::param_label(\"use_cache\")\U0060")`
`r paste0("#' @param cache_lifespan \U0060r pkgsnippets::param_label(\"cache_lifespan\")\U0060")`
`r paste0("#' @return \U0060r pkgsnippets::return_label(\"data\")\U0060")`
```

which will result in

```r
#' @param use_cache `r pkgsnippets::param_label("use_cache")`
#' @param cache_lifespan `r pkgsnippets::param_label("cache_lifespan")`
#' @return `r pkgsnippets::return_label("data")`
```

## Installation
```{r, install-snippet-dev, child = pkgsnippets::snippet_path("installation-notice_dev-version_gitlab.Rmd")}
```

## Development

### Literate Programming
```{r, literate-programming-snippet, child = pkgsnippets::snippet_path("literate-programming-notice.Rmd")}
```

### Coding Style
```{r, coding-style-snippet, child = pkgsnippets::snippet_path("coding-style-notice.Rmd")}
```
