---
title: "Introduction to pkgsnip"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to pkgsnip}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(rlang,
        include.only = c("%|%", "%||%"))
library(magrittr,
        include.only = c("%>%", "%<>%", "%T>%", "%!>%", "%$%"))
```

In the following, an overview is given of what's included in this package.

## R Markdown file snippets {#r-markdown-file-snippets}

The (R) Markdown file snippets can be used anywhere [(R) Markdown](https://rmarkdown.rstudio.com/) input is supported. For example, you can use them as input to
[knitr](https://yihui.org/knitr/)'s [`child` document option](https://yihui.org/knitr/options/#child-documents):

`r paste0("\x60\x60\x60\x60\n\x60\x60\x60", '{r, child = pkgsnip::snip_path("pkg-instl-dev-gitlab.Rmd")}\n', "\x60\x60\x60\n\x60\x60\x60\x60")`

When inserting snippets that require the name of the calling package like `"pkg-instl-dev-gitlab.Rmd"`, you should either use `pal::build_readme()` to build the
`.Rmd` file or manually assign `pkg_metadata <- pal::desc_list()` to the environment passed to
[`devtools::build_rmd()`](https://devtools.r-lib.org/reference/build_rmd.html) or its `...` argument respectively.

You can also use the snippets in [roxygen2](https://roxygen2.r-lib.org/)'s [`@includeRmd`
tag](https://roxygen2.r-lib.org/articles/rd.html#including-external--rmd-md-files):

``` rd
#' @includeRmd `r '\x60r pkgsnip::snip_path("pkgpurl.Rmd")\x60'`
```

For snippets that require the name of the calling package, just assign `pkg_metadata` (once) before inserting the snippet(s), e.g. like this:

``` rd
#' `r "\x60r pkg_metadata <- pal::desc_list()\x60"`
#'
#' @includeRmd `r '\x60r pkgsnip::snip_path("pkg-instl-dev.Rmd")\x60'`
```

### Available snippets

Currently, the following R Markdown file snippets are included:

```{r}
#| echo: false
#| results: asis

fs::path_package("snippets/",
                 package = pal::desc_value(key = "Package")) %>%
  fs::dir_ls(recurse = TRUE,
             type = "file") %>%
  fs::path_file() %>%
  paste0("- [`", ., "`](",
         stringr::str_subset(string = desc::desc_get_urls(),
                             pattern = "https://git(lab|hub)"),
         "/-/tree/master/inst/snippets/", ., ")") %>%
  pal::cat_lines()
```

## Markdown snippets

The Markdown snippets basically serve the same purpose as the *R Markdown file snippets* [above](#r-markdown-file-snippets), but are geared towards shorter
content and their semantics slightly differ. They support [glue's string interpolation](https://glue.tidyverse.org/) via curly braces, i.e. embedded R code
allows for dynamic content.

The snippets can be used in roxygen2 documentation via [inline R code](https://roxygen2.r-lib.org/articles/reuse.html#inline-code) as follows:

``` r
#' Awesome function title
#'
#' Awesome function description.
#'
#' `r "\U0060r pkgsnip::md_snip(\"rstudio_addin\")\U0060"`
```

which will result in

``` r
#' Awesome function title
#'
#' Awesome function description.
#'
#' `r pkgsnip::md_snip("rstudio_addin")`
```

### Available snippets

Currently, the following Markdown snippets are included:

```{r}
#| echo: false
#| results: asis

pkgsnip::data_md_snips |>
  pkgsnip:::backtickify_cols(cols = tidyselect::everything()) |>
  pal::pipe_table()
```

## roxygen2 tag labels

The [roxygen2](https://roxygen2.r-lib.org/) tag labels are specifically meant for [`@`-tag](https://roxygen2.r-lib.org/articles/rd.html)-based documentation
using [inline R code](https://roxygen2.r-lib.org/articles/reuse.html#inline-code).

Insert them in your functions using inline R code, e.g. as follows

``` r
`r "#' @param use_cache \U0060r pkgsnip::param_lbl(\"use_cache\")\U0060"`
`r "#' @param max_cache_age \U0060r pkgsnip::param_lbl(\"max_cache_age\")\U0060"`
`r "#' @return \U0060r pkgsnip::return_lbl(\"tibble\")\U0060"`
```

which will result in

``` r
#' @param use_cache `r pkgsnip::param_lbl("use_cache")`
#' @param max_cache_age `r pkgsnip::param_lbl("max_cache_age")`
#' @return `r pkgsnip::return_lbl("tibble")`
```

## Common abbreviations

Finally, there's a table included that lists an (opinionated) set of abbreviations commonly used when writing R code. It might come in handy as a reference, for
example, to check availability when considering using some abbreviation in a function or argument name.

The table can be retrieved by calling `pkgsnip::abbrs()` and currently looks as follows (click to expand):

```{asis}
<details><summary>Abbreviations</summary>
```

```{r}
#| echo: false
pkgsnip::abbrs() |>
  dplyr::mutate(dplyr::across(where(is.list),
                              \(x) purrr::map_chr(x,
                                                  paste0,
                                                  collapse = ", "))) |>
  knitr::kable()
```

```{asis}
</details>
```
