---
title: "Introduction to pkgsnip"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to pkgsnip}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(rlang,
        include.only = c("%|%", "%||%"))
library(magrittr,
        include.only = c("%>%", "%<>%", "%T>%", "%!>%", "%$%"))
```

In the following, an overview is given of what's included in this package.

## R Markdown file snippets

The (R) Markdown file snippets can be used anywhere [(R) Markdown](https://rmarkdown.rstudio.com/) input is supported. For example, you can use them as input to
[knitr](https://yihui.org/knitr/)'s [`child` document option](https://yihui.org/knitr/options/#child-documents):

`r paste0("\x60\x60\x60\x60\n\x60\x60\x60", '{r, child = pkgsnip::snip_path("installation-notice_dev-version_gitlab.Rmd")}\n', "\x60\x60\x60\n\x60\x60\x60\x60")`

When inserting snippets that require the name of the calling package like `"installation-notice_dev-version_gitlab.Rmd"`, you should either use
`pal::build_readme()` to build the `.Rmd` file or manually assign `pkg_metadata <- pal::desc_list()` to the environment passed to
[`devtools::build_rmd()`](https://devtools.r-lib.org/reference/build_rmd.html) or its `...` argument respectively.

You can also use the snippets in [roxygen2](https://roxygen2.r-lib.org/)'s [`@includeRmd`
tag](https://roxygen2.r-lib.org/articles/rd.html#including-external--rmd-md-files):

``` rd
#' @includeRmd `r '\x60r pkgsnip::snip_path("rmd-package-notice.Rmd")\x60'`
```

For snippets that require the name of the calling package, just assign `pkg_metadata` (once) before inserting the snippet(s), e.g. like this:

``` rd
#' `r "\x60r pkg_metadata <- pal::desc_list()\x60"`
#'
#' @includeRmd `r '\x60r pkgsnip::snip_path("installation-notice_dev-version.Rmd")\x60'`
```

### Available snippets

Currently, the following R Markdown file snippets are included:

```{r}
#| echo: false
#| results: asis

fs::path_package("snippets/",
                 package = pal::desc_value(key = "Package")) %>%
  fs::dir_ls(recurse = TRUE,
             type = "file") %>%
  fs::path_file() %>%
  paste0("- [`", ., "`](",
         stringr::str_subset(string = desc::desc_get_urls(),
                             pattern = "https://git(lab|hub)"),
         "/-/tree/master/inst/snippets/", ., ")") %>%
  pal::cat_lines()
```

## roxygen2 tag labels

The [roxygen2](https://roxygen2.r-lib.org/) tag labels are specifically meant for [`@`-tag](https://roxygen2.r-lib.org/articles/rd.html)-based documentation
using [dynamic R code](https://roxygen2.r-lib.org/articles/rd-formatting.html#dynamic-r-code-1).

Insert them in your functions using inline R code, e.g. as follows

``` r
`r paste0("#' @param use_cache \U0060r pkgsnip::param_label(\"use_cache\")\U0060")`
`r paste0("#' @param max_cache_age \U0060r pkgsnip::param_label(\"max_cache_age\")\U0060")`
`r paste0("#' @return \U0060r pkgsnip::return_label(\"data\")\U0060")`
```

which will result in

``` r
#' @param use_cache `r pkgsnip::param_label("use_cache")`
#' @param max_cache_age `r pkgsnip::param_label("max_cache_age")`
#' @return `r pkgsnip::return_label("data")`
```

## R condition messages

These are meant to be used as message texts for base R's [`message()`](https://rdrr.io/r/base/message.html), [`warning()`](https://rdrr.io/r/base/warning.html)
and [`stop()`](https://rdrr.io/r/base/stop.html) functions or [`rlang::inform()`, `rlang::warn()` and
`rlang::abort()`](https://rlang.r-lib.org/reference/abort.html) from the tidyverse respectively.

A message can be retrieved using `pkgsnip::msg()`. For some of the messages, arguments need to be provided so they can be tailored to their specific purpose.
For example, the `pkg_required` message requires the `pkg` argument:

``` r
pkgsnip::msg(name = "pkg_required",
             pkg = "some_pkg")
```

Which results in the following message text:

```{r}
#| echo: false
#| comment: ''
pkgsnip::msg(name = "pkg_required",
             this_pkg = "some_pkg")
```

Currently, the following R condition messages are included:

```{r}
#| echo: false
pkgsnip::msgs() |>
  dplyr::mutate(dplyr::across(name,
                              \(x) paste0("`", x, "`"))) |>
  pkgsnip:::add_args_col() |>
  pal::pipe_table()
```

## Common abbreviations

Finally, there's a table included that lists an (opinionated) set of abbreviations commonly used when writing R code. It might come in handy as a reference, for
example, to check availability when considering using some abbreviation in a function or argument name.

The table can be retrieved by calling `pkgsnip::abbreviations()` and currently looks as follows (click to expand):

```{asis}
<details><summary>Abbreviations</summary>
```

```{r}
#| echo: false
pkgsnip::abbreviations() |>
  dplyr::mutate(dplyr::across(where(is.list),
                              \(x) purrr::map_chr(x,
                                                  paste0,
                                                  collapse = ", "))) |>
  knitr::kable()
```

```{asis}
</details>
```
